---
title: "Clustering Analysis of Stock Returns"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Clustering Stock Returns

```{r read data}
data = read.csv("returns.csv")
returns = data[,3:122]
```

## a. Companies in Industry Sectors

The output below shows the number of companies in each industry sector.

```{r companies_in_sectors, message=FALSE, warning=FALSE}
library(dplyr)
data_byIndustry = data %>% group_by(Industry)
industry_size = data_byIndustry %>% summarise(CompanyNum=n())
as.data.frame(industry_size)
```

For each sector, plots of the average stock return over the 2008-10 period are shown in Figure 1.

```{r plot average returns, message=FALSE, warning=FALSE, fig.align='center', fig.dim=c(7,9), fig.cap='The Average Stock Return of Each Industry Secotr over the 2008-10 Period', dev.args=list(pointsize=10)}
data4plot = data[,c(2,25:60)]
avg_returns = aggregate(data4plot[,-1], by=list(Industry=data4plot$Industry), mean)
date = colnames(avg_returns[,-1])
date = sub('avg', '', date)

par(mfrow=c(5,2))
for (i in 1: dim(avg_returns)[1]){
  indus = avg_returns$Industry[i]
  indus_return = t(avg_returns[i, -1])
  plot(indus_return, type='l', xaxt='n', xlab='Month', ylab='Average Return', main=indus)
  axis(1, at=1:36, labels=as.factor(date))
}
```

**Comment:** We see from Figure 1 that patterns of average return fluctuations in each industry sector display some difference. This implies that "Industry" may be a good indicator of clustering stock returns.

## b. Hierarchical Clustering of the Average Returns

Cluster the data using hierarchical clustering, using the "Ward D2" measure of cluster dissimilarity. Figure 2 shows the cluster dendrogram of the average returns.

```{r clustering, message=FALSE, warning=FALSE, fig.align='center', fig.dim=c(7,2.5), fig.cap='Cluster Dendrogram of the Average Returns'}
library(flexclust)
d = dist(returns)
hclust.mod = hclust(d, method="ward.D2")
par(mar=c(0.5,2.5,1,0.5))
plot(hclust.mod, labels=F, ylab="Dissimilarity", xlab = "", sub = "", main="")
```

Figure 3 shows the scree plot. According to the scree plot, eight is a reasonable number of clusters for this problem because eight is a turning point.

```{r scree, fig.align='center', fig.dim=c(5,2.7), fig.cap='Scree Plot of the Average Return Clusters'}
hc.dissim = data.frame(k=seq_along(hclust.mod$height),dissimilarity=rev(hclust.mod$height))
par(mar=c(4, 4, 1, 0.5))
plot(hc.dissim$k,hc.dissim$dissimilarity,type="l",xlim=c(0,20),ylim=c(2,6.5),
     xlab="k",ylab="Dissimilarity")
axis(side=1, at=1:10)
```

## c. Extract Cluster Assignments from the Hierarchical Clustering

Compute the number of companies in each cluster from each industry sector. Results are shown below.

```{r analyze clusters, results='hold'}
h.clusters = cutree(hclust.mod, 8) # we choose k=8
data = cbind(cluster=h.clusters, data)
industry_counts = data.frame(Industry=sort(unique(data$Industry)))
for (k in 1:8){
  cluster_k = data[data$cluster == k, 1:3]
  indus_count = cluster_k %>% group_by(Industry) %>% summarise(CompanyNum = n())
  colnames(indus_count) = c("Industry", paste0("Cluster", k))
  industry_counts = merge(industry_counts, indus_count, all=TRUE)
}
industry_counts[is.na(industry_counts)] = 0
print(industry_counts[,1:5]) # print the output - part I
print(industry_counts[,c(1,6:9)]) # print the output - part II
```

The average returns by cluster in October 2008 and in March 2009 are calculated as follows.

```{r calc avg returns}
aggregate(returns, by=list(h.clusters), mean) %>% select(Group.1, avg200810, avg200903)
```

To characterize each cluster qualitatively, we plot the fluctuation patterns of average returns in each cluster over the ten years, as is shown in Figure 4.

```{r plot cluster returns, message=FALSE, warning=FALSE, fig.align='center', fig.dim=c(7,7.2), fig.cap='The Average Stock Return of Each Hierarchical Cluster over March 2006 to February 2016', dev.args=list(pointsize=10)}
cluster_ret = aggregate(returns, by=list(h.clusters), mean) %>% select(-Group.1)

par(mfrow=c(4,2))
for (i in 1:8){
  main = paste("Hierarchical Cluster", i)
  plot(t(cluster_ret[i,]),type='l',xaxt='n',xlab='Month',ylab='Average Return',main=main)
  axis(1, at=seq(1, 120, 12), labels=paste0(2006:2015, "03"))
}
```

From Figure 4, we see that each cluster has the following characteristics:

+ **Cluster1:** There are 183 companies in Cluster 1, accounting for 42% of all the 433 companies. The fluctuation pattern in Cluster 1 can represent the market fluctuation.
+ **Cluster2:** Large fluctuations over the 2008-11 period, and relatively small fluctuations around 0 in other periods.
+ **Cluster3:** Large fluctuations below 0 in most periods, implying that companies in Cluster 3 do not perform well in the market.
+ **Cluster4:** Large fluctuations over the 2008-11 period. Relatively small fluctuations around 0 in other periods, but they are slightly greater than those in Cluster 2. Energy companies take up 65% of all the companies in this cluster.
+ **Cluster5:** Cluster 5 only has one financial company, whose average returns over the 2008-09 period largely fluctuated above 0. In the middle of 2009, the average return boosted to over 2.0, implying that this company gained significant benefits in the post financial crisis period. In periods other than 2008-09, the average return of the company slightly fluctuates around 0.
+ **Cluster6:** Large fluctuations over the 2008-11 period, and financial companies take up 76% of all the companies in this cluster.
+ **Cluster7:** Companies in this cluster also gained significant benefits in the post financial crisis period, but the benefits were less than that of the company in Cluster 5.
+ **Cluster8:** Large fluctuations ranging from -0.4 to 0.6 over the 2008-09 period, and still large fluctuations ranging from -0.2 to 0.2 after 2009.

## d. K-Means Clustering of the Average Returns

Cluster the data using the k-means clustering algorithm, using the same number of clusters as in Question c (k=8). After clustering the average returns by k-means, we compute the number of companies in each cluster from each industry sector. Results are shown below.

```{r KMeans, results='hold'}
set.seed(144)
km = kmeans(returns, centers = 8, iter.max=100)
km.clusters = km$cluster

# compute the number of companies in each cluster from each industry sector
data = cbind(cluster_km=km.clusters, data)
industry_counts = data.frame(Industry=sort(unique(data$Industry)))
for (k in 1:8){
  cluster_k = data[data$cluster_km == k, 1:4]
  indus_count = cluster_k %>% group_by(Industry) %>% summarise(CompanyNum = n())
  colnames(indus_count) = c("Industry", paste0("Cluster", k))
  industry_counts = merge(industry_counts, indus_count, all=TRUE)
}
industry_counts[is.na(industry_counts)] = 0

print(industry_counts[,1:5]) # print the output - part I
print(industry_counts[,c(1,6:9)]) # print the output - part II
```

Sum of the number of companies in each cluster:

```{r cluster sum}
apply(industry_counts[,-1], 2, sum)
```


We also plot the fluctuation patterns of average returns in each k-means cluster over the ten years, as is shown in Figure 5.

```{r plot KCluster returns, message=FALSE, warning=FALSE, fig.align='center', fig.dim=c(7,7), fig.cap='The Average Stock Return of Each K-Means Cluster over March 2006 to February 2016', dev.args=list(pointsize=10)}
cluster_ret = aggregate(returns, by=list(km.clusters), mean) %>% select(-Group.1)

par(mfrow=c(4,2))
for (i in 1:8){
  main = paste("K-Means Cluster", i)
  plot(t(cluster_ret[i,]),type='l',xaxt='n',xlab='Month',ylab='Average Return',main=main)
  axis(1, at=seq(1, 120, 12), labels=paste0(2006:2015, "03"))
}
```

Compare hierarchical clustering with k-means clustering:

+ Some clustering results of the two models are similar. For example, Cluster 1 of hierarchical clustering and Cluster 8 of k-means clustering both have the largest number of companies and present similar fluctuating patterns. These two clusters can represent the market fluctuation. Cluster 5 of hierarchical clustering and Cluster 6 of k-means clustering both have only one or two financial companies, and these companies gained significant benefits in the post financial crisis period.
+ In terms of similarities between the two models, both models give interpretable results for stock clusters regarding the fluctuation of average returns. We can analyze these results to inform investment strategies.
+ In terms of differences between the two models, hierarchical clustering gives a dendrogram of all the clusters, and we use a scree plot to choose a reasonable cluster number; k-means clustering runs according to a given cluster number, and this number is usually hard to decide.

## e. Use the Models to Inform Investment Strategies

The clustering models show the results of stock clusters. Each stock cluster has stocks that present similar fluctuation patterns of average returns. Different stock clusters, possessing heterogeneous fluctuating patterns, cater to different types of investors. Therefore, investors can choose stocks in certain clusters according to their personalized investment strategies. For example, risk-averse investors can choose the stock cluster that has minor fluctuations around 0; risk-seeking investors can choose the stock cluster that has significant fluctuations. The clustering models serve as a good reference for personalizing different investors' investment strategies.


